---
title: "Final Project"
author: "Sunnie Oh, InWoong Bae"
          
output: 
    pdf_document:
          fig_height: 4
          fig_width: 4
          latex_engine: xelatex
    includes:
          in_header: latex/header.tex
          before_body: latex/before_body.tex
          after_body: latex/after_body.tex
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
concrete<-read.table('Concrete.txt')
summary(concrete)
names(concrete)
head(concrete)
```

```{r}
mod.full<-lm(Y~.,data=concrete)
anova(mod.full)
```
According to the anova table above, 1021=n-8-1. So n=1030

a) forward selection with BIC
```{r}
mod.0<-lm(Y~1, data=concrete) #linear model with only intercepts
mod.full<-lm(Y~.,data=concrete) #full model
step(mod.0, scope = list(lower = mod.0, upper = mod.full), direction = 'forward', k = log(1030))
```

Do diagnostic checks 
```{r}
mod.for<-lm(Y ~ X1 + X5 + X8 + X2 + X4 + X3, data = concrete)
plot(mod.for)
```

According to the plots, model by forward selection pretty fufills noramlity, linearity, and constant variance.

Use influenceIndexPlot to find influential points to remove
```{r}
library(car)
infIndexPlot(mod.for)
```

```{r}
predict(mod.for, data.frame(X1= 200, X5=10, X8=100, X2=150, X4=180, X3=85),
interval = 'confidence', level = 0.95)
```
We are 95% confident that true mean response is between 42.126 and 44.628
```{r}
predict(mod.for, data.frame(X1= 200, X5=10, X8=100, X2=150, X4=180, X3=85),
interval = 'prediction', level = 0.95)
```
We are 95% confident that concrete compressive strength for and individual value of each predictor values is between 22.912 and 63.842

(b) Backward elimination with BIC
```{r}
step(mod.full, scope = list(lower = mod.0, upper = mod.full), direction = 'backward', k = log(1030))
```

Do diagnostic checks
```{r}
mod.back<-lm(Y ~ X1 + X2 + X3 + X4 + X5 + X8, data=concrete)
plot(mod.back)
```

According to the plots, model by forward selection pretty fufills noramlity, linearity, and constant variance.


Use influenceIndexPlot to find influential points to remove
```{r}
library(car)
infIndexPlot(mod.back)
```

```{r}
anova(mod.for, mod.back)
```

The models derived by two differnt methods are same except for the order of predictors. 
And the result of ANOVA also same.
